---
title: "TS"
author: "Dmitry Zotikov"
date: "March 19, 2017"
output: html_document
---

Earthquakes-count series.

 * Not periodic (`stl`, `decompose`) give an error.
 * SMA Smoothing shows there are two periods starting from year 1925 (but the two above functions still fail to identify any)
 * 0 ACF Forecasting:
   * Holt-Winters (simple) exponential smoothing
   * H-W ES on smoothed data
   * Confirmation with residual analysis and Box-Ljung (independence) test
 * Non-stationary -> Stationary TS by differences (1-st order)
   * Confirmation with `kpss.test` (level/trend stationarity)
 * ARIMA(0, 1, 1) forecasting

<!-- ## Comparison -->

```{r, eval=FALSE, include=FALSE}
fc.hw.s <- forecast(HoltWinters(tt, beta=FALSE, gamma=FALSE))
fc.hw <- forecast(HoltWinters(tt, gamma=FALSE))
fc.arima.011 <- forecast(arima(tt, order=c(0, 1, 1)))
hmm.eq <- PoisHmm(3, tt)
hmm.eq$fit(tt)
fc.hmm <- hmm.eq$forecast(tt, seq(min(tt), max(tt)), 10)
par(mfrow=c(1,3))
plot(fc.hw)
plot(fc.hw.s)
plot(fc.arima.011)
par(mfrow=c(1,1))
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

https://a-little-book-of-r-for-time-series.readthedocs.io/en/latest/src/timeseries.html

```{r}
source('import_earthquakes.R')
source('pois_hmm.R')
source('norm_hmm.R')
```

## ACF, tt object

```{r}
xx <- earthquakes.usgs.annual$count
plot(xx, type='b')
acf(xx)

xx.c <- xx - mean(xx)
acf(xx.c)
plot(xx.c, type='b')

tt <- ts(xx, start=earthquakes.usgs.annual$year[1])
plot(tt)
```

const noise & const amplitude => additive model
if not -- apply log

## Decomposition

### SMA Smoothing

```{r, fig.height=10, fig.width=9}
library('TTR')

par(mfrow=c(4,4))
for (n in 1:16) {
  plot(SMA(tt, n), main=n)
}
for (n in 1:16) {
  plot(SMA(log(tt), n), main=n)
}
par(mfrow=c(1,1))
```

### MA Decomposition

```{r}
plot(log(tt))
```

```{r, fig.height=10, fig.width=9}
# plot(decompose(tt)) ## nope
# plot(decompose(tt), 'multiplicative') ## nope
# stl(tt, 'periodic') ## nope
```

Periods starting from year 1924.

```{r}
plot(SMA(tt, 10), main=10)
points(SMA(tt, 10))
# identify(earthquakes.usgs.annual$year, earthquakes.usgs.annual$count)
```

Still no.

```{r}
tt.per.smoothed <- na.omit(SMA(ts(tt[(25-10):length(tt)]), 10))
plot(tt.per.smoothed)
# plot(decompose(tt.per.smoothed ))  # nope
```


## Forecasting

Exponential smoothing methods make no assumptions about the correlations between successive values of the time series.

### Simple Exponential Smoothing (Holt-Winters)

```{r}
tt.hw.s <- HoltWinters(tt, beta=FALSE, gamma=FALSE)
plot(tt.hw.s)
summary(tt.hw.s)
str(tt.hw.s)
```

Equivalent for this TS:

```{r}
library('forecast')
tt.ets <- ets(tt)  # A/N/N (error/trend/season)
```

```{r}
plot(forecast(tt.hw.s))
plot(forecast(tt.ets))
```

### Not So Simple

 * $\alpha$ --- estimate of the level at the current time point;
 * $\beta$ --- estimate of the slope b of the trend component at the current time point.

Values that are close to 0 mean that little weight is placed on the most recent observations when making forecasts of future values.

```{r}
tt.hw <- HoltWinters(tt, gamma=FALSE)
plot(forecast(tt.hw))
```

### Smoothed

```{r}
plot(forecast(HoltWinters(na.omit(SMA(tt, 10)), beta=FALSE, gamma=FALSE)))
```

### Residual analysis

The model cannot be improved upon if there's:

 * no correlations between residuals
 * residuals ~ $\mathrm{N}(0, \sigma^2)$
 * Box-Ljung doesn't reject H0: indep

```{r}
rr.s <- na.omit(forecast(tt.hw.s)$residuals)
rr <- na.omit(forecast(tt.hw)$residuals)
acf(rr.s)
acf(rr)
```


```{r}
for (resid in list(rr, rr.s)) {
  hist(resid, freq=FALSE)
  phi <- function(x) dnorm(x, 0, sd(resid))
  curve(phi, from=min(resid), to=max(resid), col='red', add=TRUE)
  qqnorm(resid)
  qqline(resid)
  print(shapiro.test(resid))
  print(Box.test(resid, lag=15, type='Ljung-Box'))
}
```



## WWWusage 

```{r}
plot(WWWusage)
```

### SMA Forecast

```{r}
plot(forecast(WWWusage))
```

### HMM

```{r}
# selectModelIc(PoisHmm, WWWusage, 10, TRUE)  # (3, 5)
hmm.www <- PoisHmm(5, WWWusage)
hmm.www$fit(WWWusage)
library('corrplot')
```

```{r}
color.bar <- function(lut, min, max=-min, nticks=11, ticks=seq(min, max, len=nticks), title='') {
  scale = (length(lut)-1)/(max-min)
  
  # dev.new(width=1.75, height=5)
  plot(c(0,10), c(min,max), type='n', bty='n', xaxt='n', xlab='', yaxt='n', ylab='', main=title)
  axis(2, ticks, las=1)
  for (i in 1:(length(lut)-1)) {
    y = (i-1)/scale + min
    rect(0,y,10,y+1/scale, col=lut[i], border=NA)
  }
}
col2 <- colorRampPalette(c("#67001F", "#B2182B", "#D6604D", "#F4A582", "#FDDBC7",
                           "#FFFFFF", "#D1E5F0", "#92C5DE", "#4393C3", "#2166AC", "#053061"))
```

```{r}
corrplot(hmm.www$A, method='circle', is.corr=FALSE, addCoef.col = 'black',
         col=col2(100)[30:100])
```

```{r, eval=FALSE, fig.height=10, fig.width=9, include=FALSE}
www.probes <- seq(min(WWWusage), max(WWWusage), by=3)
www.probes <- seq(150, 350, by=4)
www.fc <- hmm.www$forecast(WWWusage, www.probes, 16)
hmm.www$plotForecast(www.fc, WWWusage, www.probes, 1:16)
```

## ARIMA

Defined for stationary series.

### Differences

Not stationary in mean (use HMM algos to detect component means?)

```{r}
par(mfrow=c(2,2))
plot(tt)
for (n in 1:3) {
  plot(diff(tt, differences=n))
}
par(mfrow=c(1,1))
```

```{r}
tt.diff <- diff(tt, differences=1)
acf(tt.diff)
```

Box-Ljung rejected => non-independent?

```{r}
print(Box.test(tt.diff, lag=15, type='Ljung'))
```

H0: level/trend stationary

```{r}
library('tseries')
# adf.test(tt.diff, alternative='stationary')
print(kpss.test(tt))
print(kpss.test(tt.diff))
```

### Selecting ARIMA model

`which(abs(acf(tt.diff, plot=FALSE)$acf))` = 1, 2 so ARMA(0, 1) for `tt.diff` and
ARIMA(0, 1, 1) for tt:

```{r}
print(auto.arima(tt))
```

### Forecasting

```{r}
plot(forecast(arima(tt, order=c(0, 1, 1))))
```

+ residual analysis
